PrologPF is a parallelising compiler targeting a distributed system of
general purpose workstations connected by a relatively low performance
network.  The source language extends standard Prolog with the
integration of higher-order functions.

The execution of a compiled PrologPF program proceeds in a similar
manner to standard Prolog, but uses \textit{oracles} in one of two
modes.  An oracle represents the sequence of clauses used to reach a
given point in the problem search tree, and the same PrologPF executable
can be used to \textit{build} oracles, or \textit{follow} oracles
previously generated.

The parallelisation strategy used by PrologPF proceeds in two phases,
which this research shows can be interleaved.
An initial phase searches the problem tree to a limited depth,
recording the discovered incomplete paths.
In the second phase these paths are allocated
to the available processors in the network.  Each processor
follows its assigned paths and fully searches the referenced subtree,
sending solutions back to a control processor.  This research
investigates the use of the technique with a one-time partitioning
of the problem and no further scheduling communication, and with the
recursive application of the partitioning technique to effect
dynamic work reassignment.

For a problem requiring \textit{all} solutions to be found, execution
completes when all the distributed processors have completed the
search of their assigned subtrees.  If \textit{one} solution is
required, the execution of all the path processors is terminated when
the control processor receives the first solution.

The presence of the extra-logical Prolog predicate \textit{cut} in the
user program conflicts with the use of oracles to represent valid open
subtrees.  PrologPF promotes the use of higher-order functional
programming as an alternative to the use of \textit{cut}.  The
combined language shows that functional support can be added as a
consistent extension to standard Prolog.

